Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-30 paper.qmd
Generate embeddings with 1000 points for Five Gaussian clusters
Fixed the issue of having model and data little bit seprated than usual

# Function to scale data manually
scale_data_manual <- function(data) {
  # Step 1: Center the data (mean 0)
  data_centered <- center_data(data)

@@

@@

# Apply the scaling function to the s_curve_noise data
scaled_s_curve_noise <- scale_data_manual(
  training_data_scurve |> select(-ID)) |>
  as_tibble()

# Scaling the true model and add the mean of scaled noise dimensions
scaled_s_curve_true <- scale_data_manual(
  true_model_df |> 
    select(-ID) |>
    select(x1:x3)) |>
  as_tibble() |>
  mutate(x4 = mean(scaled_s_curve_noise$x4),
         x5 = mean(scaled_s_curve_noise$x5),
         x6 = mean(scaled_s_curve_noise$x6),
         x7 = mean(scaled_s_curve_noise$x7))

@@

training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau_training.rds")

tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_61.rds")
gau1_scaled_obj <- gen_scaled_data(
  data = tsne_data_gau)
tsne_gau_scaled <- gau1_scaled_obj$scaled_nldr

@@

# Apply the scaling
scaled_gau_data <- scale_data_manual(training_data_gau |> select(-ID)) |>
  as_tibble()

@@

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
  dplyr::select(-hb_id) |>
  select(-type)

# Apply the scaling
scaled_gau_data_model <- scale_data_manual(df_b) |>
  as_tibble()

@@

## Second projection
projection <- cbind(
    c(-0.71895,0.29250,-0.29905,-0.01656),
    c(-0.16698,-0.62707,-0.23756,0.46328))

@@

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
  dplyr::select(-hb_id)|>
  select(-type)

# Apply the scaling
scaled_gau_data_model <- scale_data_manual(df_b) |>
  as_tibble()

@@

# Apply the scaling
scaled_gau_data_model <- scale_data_manual(df_b) |>
  as_tibble()

@@

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
  dplyr::select(-hb_id)|>
  select(-type)

# Apply the scaling
scaled_gau_data_model <- scale_data_manual(df_b) |>
  as_tibble()

@@

### Second projection
projection <- cbind(
    c(-0.60995,0.33361,0.32920,-0.31518),
    c(0.19525,-0.30554,0.74417,0.07600))

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  limits = 0.3,
  axis_pos_x = -0.05,
  axis_pos_y = -0.4)

@@

five_gau_proj_pacmap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.8,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.2)) +
  ylim(c(-0.45, 0)) +
  interior_annotation("c2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

# Apply the scaling
scaled_pbmc_data <- scale_data_manual(training_data_pbmc |> select(-ID)) |>
  as_tibble()

@@

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
  dplyr::select(-hb_id) |>
  select(-type)

# Apply the scaling
scaled_pbmc_data_model <- scale_data_manual(df_b_pbmc) |>
  as_tibble()

@@

# Apply the scaling
scaled_pbmc_data <- scale_data_manual(training_data_pbmc |> select(-ID)) |>
  as_tibble()

@@

@@

@@

<!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
```{r}
#| label: mnist-model-proj

# Apply the scaling
scaled_mnist_data <- scale_data_manual(training_data_mnist |> select(-ID)) |>
  as_tibble()

df_b_mnist <- df_bin_mnist |>
  dplyr::filter(hb_id %in% df_bin_centroids_mnist$hexID) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b_mnist <- df_b_mnist[match(df_bin_centroids_mnist$hexID, df_b_mnist$hb_id),] |>
  dplyr::select(-hb_id) |>
  select(-type)

# Apply the scaling
scaled_mnist_data_model <- scale_data_manual(df_b_mnist) |>
  as_tibble()

>>>>

# Function to scale data manually
scale_data_manual <- function(data, type_col) {
  # Step 1: Center the data (mean 0)
  data_centered <- center_data(data |> select(-all_of(type_col)))

@@

  # Combine the scaled data with the 'type' column and return as a tibble
  data_scaled <- as_tibble(data_scaled) %>% 
    mutate(!!type_col := data[[type_col]])

@@

data_scurve <- training_data_scurve |> 
  select(-ID) |> 
  mutate(type = "data")

model_scurve <- true_model_df |> 
    select(-ID) |> 
  mutate(type = "model")

# Apply the scaling
df_model_data_scurve <- bind_rows(data_scurve, model_scurve)
scaled_scurve <- scale_data_manual(df_model_data_scurve, "type") |>
  as_tibble()

scaled_s_curve_noise <- scaled_scurve |>
  filter(type == "data") |>
  select(-type)

scaled_s_curve_true <- scaled_scurve |>
  filter(type == "model") |>
  select(-type)

@@

training_data_gau <- read_rds("data/five_gau_clusters/data_five_gau.rds")

data_gau <- training_data_gau |> 
  select(-ID) |>
  mutate(type = "data")

tsne_data_gau <- read_rds("data/five_gau_clusters/tsne_data_five_gau_71.rds")
gau1_scaled_obj <- gen_scaled_data(
  data = tsne_data_gau)
tsne_gau_scaled <- gau1_scaled_obj$scaled_nldr

@@

@@

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
  dplyr::select(-hb_id)

# Apply the scaling
df_model_data <- bind_rows(data_gau, df_b)
scaled_gau <- scale_data_manual(df_model_data, "type") |>
  as_tibble()

scaled_gau_data <- scaled_gau |>
  filter(type == "data") |>
  select(-type)

scaled_gau_data_model <- scaled_gau |>
  filter(type == "model") |>
  select(-type)

@@

## Second projection
# projection <- cbind(
#     c(-0.71895,0.29250,-0.29905,-0.01656),
#     c(-0.16698,-0.62707,-0.23756,0.46328))
projection <- cbind(
  c(0.36030,0.00630,0.66545,-0.34307),
  c(-0.01861,0.14409,-0.36798,-0.73066))

@@

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
  dplyr::select(-hb_id)

@@

# Apply the scaling
df_model_data <- bind_rows(data_gau, df_b)
scaled_gau <- scale_data_manual(df_model_data, "type") |>
  as_tibble()

scaled_gau_data <- scaled_gau |>
  filter(type == "data") |>
  select(-type)

scaled_gau_data_model <- scaled_gau |>
  filter(type == "model") |>
  select(-type)

@@

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b <- df_b[match(df_bin_centroids_gau1$hexID, df_b$hb_id),] |>
  dplyr::select(-hb_id)

# Apply the scaling
df_model_data <- bind_rows(data_gau, df_b)
scaled_gau <- scale_data_manual(df_model_data, "type") |>
  as_tibble()

scaled_gau_data <- scaled_gau |>
  filter(type == "data") |>
  select(-type)

scaled_gau_data_model <- scaled_gau |>
  filter(type == "model") |>
  select(-type)

@@

### Second projection
# projection <- cbind(
#     c(-0.60995,0.33361,0.32920,-0.31518),
#     c(0.19525,-0.30554,0.74417,0.07600))
projection <- cbind(
  c(0.36030,0.00630,0.66545,-0.34307),
  c(-0.01861,0.14409,-0.36798,-0.73066))

@@

axes_obj <- gen_axes(
  data = scaled_gau_data,
  proj_data = projected_df,
  limits = 0.3,
  axis_pos_x = -0.7,
  axis_pos_y = -0.7)

@@

five_gau_proj_pacmap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.8,
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.75, -0.35)) +
  ylim(c(-0.75, -0.4))+
  interior_annotation("c2", 
                      position = c(0.08, 0.9),
                      cex = 2)

@@

data_pbmc <- training_data_pbmc |> 
  select(-ID) |>
  mutate(type = "data")

@@

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b_pbmc <- df_b_pbmc[match(df_bin_centroids_pbmc$hexID, df_b_pbmc$hb_id),] |>
  dplyr::select(-hb_id)

# Apply the scaling
df_model_data_pbmc <- bind_rows(data_pbmc, df_b_pbmc)
scaled_pbmc <- scale_data_manual(df_model_data_pbmc, "type") |>
  as_tibble()

scaled_pbmc_data <- scaled_pbmc |>
  filter(type == "data") |>
  select(-type)

scaled_pbmc_data_model <- scaled_pbmc |>
  filter(type == "model") |>
  select(-type)

@@

@@

<!--add different NLDR layouts-->

@@

data_mnist <- training_data_mnist |>
  select(-ID) |>
  mutate(type = "data")

@@

<!--bin1 = 22, bin2 = 17, b = 374, non_empty = 140-->
```{r}
#| label: mnist-model-proj
df_b_mnist <- df_bin_mnist |>
  dplyr::filter(hb_id %in% df_bin_centroids_mnist$hexID) |>
  dplyr::mutate(type = "model") ## Data with summarized mean

## Reorder the rows of df_b according to the hexID order in df_b_with_center_data
df_b_mnist <- df_b_mnist[match(df_bin_centroids_mnist$hexID, df_b_mnist$hb_id),] |>
  dplyr::select(-hb_id)

# Apply the scaling
df_model_data_mnist <- bind_rows(data_mnist, df_b_mnist)
scaled_mnist <- scale_data_manual(df_model_data_mnist, "type") |>
  as_tibble()

scaled_mnist_data <- scaled_mnist |>
  filter(type == "data") |>
  select(-type)

scaled_mnist_data_model <- scaled_mnist |>
  filter(type == "model") |>
  select(-type)
